---
title: "Honda Clustering"
author: "Raymond David"
date: "2023-11-21"
output:
  pdf_document: default
  html_document: default
---


```{r include=FALSE}
# Load Libraries
library(tidyverse)
```


# Clustering
```{r}
features <- read_csv("feature.csv")
merged <- read_csv("merged.csv")
```




```{r}
library(stats)

# Function to extract numeric values
extract_numeric <- function(x) {
  as.numeric(gsub("[^0-9.]", "", x))
}

# Select columns for clustering
selected_columns <- data.frame(
  Acceleration = extract_numeric(features$Acceleration),
  TopSpeed = extract_numeric(features$TopSpeed),
  Range = extract_numeric(features$Range),
  Efficiency = extract_numeric(gsub("[^0-9.]", "", features$Efficiency)),
  FastChargeSpeed = extract_numeric(features$FastChargeSpeed)
)

# Normalize the data
normalized_data <- scale(selected_columns)

# Determine the number of clusters (k value)
# For demonstration purposes, let's assume k = 3
k <- 3

# Perform k-means clustering
kmeans_result <- kmeans(normalized_data, centers = k)

# View the cluster assignments
cluster_assignments <- kmeans_result$cluster
#print(cluster_assignments)

# View the centroids of each cluster
centroids <- kmeans_result$centers
print(centroids)
```
**Interpreation**
\
Cluster 1:

Acceleration: The centroid value is close to zero but slightly negative, suggesting that vehicles in this cluster tend to have slightly lower than average acceleration.
TopSpeed: Similar to acceleration, the value is close to zero but slightly negative, indicating that vehicles in this cluster might have slightly lower than average top speeds.
Range: Again, close to zero but slightly negative, implying that vehicles in this cluster might have slightly lower than average range capabilities.
Efficiency: The value is positive, suggesting that these vehicles might have slightly higher efficiency compared to the dataset's average.
FastChargeSpeed: Positive centroid value suggests above-average fast charge speeds, though not significantly higher compared to other clusters.
Cluster 2:

Acceleration: The value is positive, indicating that vehicles in this cluster tend to have higher than average acceleration.
TopSpeed: Negative centroid indicates that vehicles in this cluster have lower than average top speeds.
Range: Negative centroid suggests that vehicles in this cluster might have lower than average range capabilities.
Efficiency: Negative centroid implies lower efficiency compared to the dataset's average.
FastChargeSpeed: Negative centroid suggests below-average fast charge speeds.
Cluster 3:

Acceleration: Negative centroid, implying lower than average acceleration for vehicles in this cluster.
TopSpeed: Positive centroid indicates that vehicles in this cluster might have higher than average top speeds.
Range: Positive centroid suggests that vehicles in this cluster might have higher than average range capabilities.
Efficiency: Negative centroid implies lower efficiency compared to the dataset's average.
FastChargeSpeed: Positive centroid suggests above-average fast charge speeds.
These interpretations provide insights into the average representation of different features within each cluster formed by the k-means clustering algorithm. Each cluster represents a group of observations with specific characteristics regarding performance attributes, efficiency, and fast charge capabilities.

Based on these centroids, you can interpret the characteristics of each cluster:

Cluster 1 seems to represent vehicles with moderate acceleration, slightly lower than average top speed, range, but higher efficiency and above-average fast charge speed.
Cluster 2 represents vehicles with lower than average acceleration, top speed, range, and efficiency, along with below-average fast charge speed.
Cluster 3 represents vehicles with the highest top speed, range, but lower efficiency and the highest fast charge speed.


```{r}
# Assuming kmeans_result is your kmeans clustering result
plot(normalized_data, col = kmeans_result$cluster, pch = 19)
points(kmeans_result$centers, col = 1:k, pch = 8, cex = 2)

```





```{r}
# K-means + One Hot Encoding
# Extract numerical values from columns
extract_numeric <- function(x) {
  as.numeric(gsub("[^0-9.]", "", x))
}

# Select numerical columns for clustering
selected_columns <- data.frame(
  Acceleration = extract_numeric(features$Acceleration),
  TopSpeed = extract_numeric(features$TopSpeed),
  Range = extract_numeric(features$Range),
  Efficiency = extract_numeric(gsub("[^0-9.]", "", features$Efficiency)),
  FastChargeSpeed = extract_numeric(features$FastChargeSpeed)
)

# One-hot encode categorical column 'Drive'
drive_column <- model.matrix(~ Drive - 1, data = features)
selected_columns <- cbind(selected_columns, drive_column)

# Normalize the data
normalized_data <- scale(selected_columns)

# Determine the number of clusters (k value)
# For demonstration purposes, let's assume k = 3
k <- 3

# Perform k-means clustering
kmeans_result <- kmeans(normalized_data, centers = k)

# View the cluster assignments
cluster_assignments <- kmeans_result$cluster
#print(cluster_assignments)

# View the centroids of each cluster
centroids <- kmeans_result$centers
print(centroids)

```

**Interpreation**
\
Cluster 1:

Acceleration: The value is negative, indicating that vehicles in this cluster tend to have lower than average acceleration.
TopSpeed: The value is positive, suggesting that vehicles in this cluster have higher than average top speeds.
Range: The value is positive, indicating that vehicles in this cluster tend to have higher than average range capabilities.
Efficiency: The value is positive, suggesting that these vehicles might have higher efficiency compared to the dataset's average.
FastChargeSpeed: Positive value implies above-average fast charge speeds.
DriveAll Wheel Drive: The centroid is close to 1, indicating a strong representation of All-Wheel Drive vehicles in this cluster.
DriveFront Wheel Drive & DriveRear Wheel Drive: Both have negative centroid values, indicating fewer occurrences or a lesser representation of Front-Wheel Drive and Rear-Wheel Drive vehicles compared to the average.
Cluster 2:

Acceleration: The value is positive, suggesting that vehicles in this cluster tend to have higher than average acceleration.
TopSpeed: Negative centroid indicates that vehicles in this cluster have lower than average top speeds.
Range: Negative centroid suggests that vehicles in this cluster might have lower than average range capabilities.
Efficiency: Negative centroid implies lower efficiency compared to the dataset's average.
FastChargeSpeed: Positive centroid suggests above-average fast charge speeds.
DriveAll Wheel Drive: The centroid is close to -1, indicating a strong representation of All-Wheel Drive vehicles in this cluster.
DriveRear Wheel Drive: Positive centroid value indicates a stronger representation of Rear-Wheel Drive in this cluster compared to the average.
Cluster 3:

Acceleration: Positive centroid, implying higher than average acceleration for vehicles in this cluster.
TopSpeed: Negative centroid indicates that vehicles in this cluster might have lower than average top speeds.
Range: Negative centroid suggests that vehicles in this cluster might have lower than average range capabilities.
Efficiency: Negative centroid implies lower efficiency compared to the dataset's average.
FastChargeSpeed: Negative centroid suggests below-average fast charge speeds.
DriveAll Wheel Drive & DriveFront Wheel Drive: Both have negative centroid values, indicating fewer occurrences or a lesser representation of All-Wheel Drive and Front-Wheel Drive vehicles compared to the average. However, Front-Wheel Drive is relatively stronger in this cluster compared to the average.

Each cluster represents a distinct group of vehicles with different characteristics in terms of performance, drivetrain types, efficiency, and fast charge capabilities.
Cluster 1 seems to include high-range, efficient vehicles with higher top speeds, predominantly All-Wheel Drive.
Cluster 2 consists of vehicles with lower range, efficiency, and top speeds, featuring above-average acceleration and varying distributions of drivetrain types.
Cluster 3 includes vehicles with higher acceleration but lower range, efficiency, and top speeds, with differing distributions of drivetrain types compared to the dataset's average.




# Cluster using Merged Dataset

```{r}
# Extract numerical values from columns
extract_numeric <- function(x) {
  as.numeric(gsub("[^0-9.]", "", x))
}

# Select columns for clustering
selected_columns <- data.frame(
  SalesCount = merged$`Sales Count`,
  Acceleration = extract_numeric(merged$`Acceleration (sec)`),
  TopSpeed = extract_numeric(merged$`TopSpeed (km/h)`),
  Range = extract_numeric(merged$`Range (km)`)
)

# Normalize the data
normalized_data <- scale(selected_columns)

# Determine the number of clusters (k value)
# For demonstration purposes, let's assume k = 3
k <- 3

# Perform k-means clustering
# Convert data.frame to matrix as kmeans() requires a matrix input
kmeans_result <- kmeans(as.matrix(normalized_data), centers = k)

# View the cluster assignments
cluster_assignments <- kmeans_result$cluster
print(cluster_assignments)

# View the centroids of each cluster
centroids <- kmeans_result$centers
print(centroids)
```
**Interpreation**
\
Cluster 1:

SalesCount: The centroid value is positive, indicating that vehicles in this cluster tend to have a higher than average sales count.
Acceleration: The value is positive, suggesting higher than average acceleration for vehicles in this cluster.
TopSpeed: The value is negative, indicating a lower than average top speed for vehicles in this cluster.
Range: The value is negative, suggesting vehicles in this cluster tend to have a lower than average range.
Cluster 2:

SalesCount: The centroid value is negative, suggesting that vehicles in this cluster tend to have a lower than average sales count.
Acceleration: The value is positive, implying higher than average acceleration for vehicles in this cluster.
TopSpeed: The value is negative, suggesting a lower than average top speed for vehicles in this cluster.
Range: The value is negative, indicating vehicles in this cluster tend to have a lower than average range.
Cluster 3:

SalesCount: The centroid value is positive, indicating that vehicles in this cluster tend to have a higher than average sales count.
Acceleration: The value is negative, suggesting lower than average acceleration for vehicles in this cluster.
TopSpeed: The value is positive, indicating a higher than average top speed for vehicles in this cluster.
Range: The value is positive, suggesting vehicles in this cluster tend to have a higher than average range.
```{r}
# K-means + One-hot Encoding
# Extract numerical values from columns
extract_numeric <- function(x) {
  as.numeric(gsub("[^0-9.]", "", x))
}

# Manually encode 'BodyStyle' column
encoded_body_style <- model.matrix(~ `Body Style` - 1, data = merged)

# Select numerical columns for clustering
selected_columns <- cbind(
  encoded_body_style,
  SalesCount = merged$`Sales Count`,
  Acceleration = extract_numeric(merged$`Acceleration (sec)`),
  TopSpeed = extract_numeric(merged$`TopSpeed (km/h)`)
)

# Normalize the data
normalized_data <- scale(selected_columns)

# Determine the number of clusters (k value)
# For demonstration purposes, let's assume k = 3
k <- 3

# Perform k-means clustering
kmeans_result <- kmeans(normalized_data, centers = k)

# View the cluster assignments
cluster_assignments <- kmeans_result$cluster
# print(cluster_assignments)

# View the centroids of each cluster
centroids <- kmeans_result$centers
print(centroids)
```
**Interpretation**
\
Let's interpret these centroids:

Cluster 1:

Body Style (Hatchback): The centroid value is positive, indicating a stronger representation of hatchback vehicles in this cluster.
Body Style (Sedan): The value is negative, suggesting fewer occurrences or a lesser representation of sedan vehicles in this cluster.
Body Style (SUV): The value is negative, indicating fewer occurrences or a lesser representation of SUV vehicles in this cluster.
SalesCount: The value is close to zero, suggesting an average sales count for vehicles in this cluster.
Acceleration: The value is positive, implying higher than average acceleration for vehicles in this cluster.
TopSpeed: The value is negative, indicating a lower than average top speed for vehicles in this cluster.
Cluster 2:

Body Style (Hatchback): The value is negative, suggesting fewer occurrences or a lesser representation of hatchback vehicles in this cluster.
Body Style (Sedan): The centroid value is positive, indicating a stronger representation of sedan vehicles in this cluster.
Body Style (SUV): The value is negative, indicating fewer occurrences or a lesser representation of SUV vehicles in this cluster.
SalesCount: The value is positive, suggesting a higher than average sales count for vehicles in this cluster.
Acceleration: The value is negative, implying lower than average acceleration for vehicles in this cluster.
TopSpeed: The value is positive, suggesting a higher than average top speed for vehicles in this cluster.
Cluster 3:

Body Style (Hatchback): The value is negative, suggesting fewer occurrences or a lesser representation of hatchback vehicles in this cluster.
Body Style (Sedan): The value is negative, suggesting fewer occurrences or a lesser representation of sedan vehicles in this cluster.
Body Style (SUV): The centroid value is positive, indicating a stronger representation of SUV vehicles in this cluster.
SalesCount: The value is close to zero, suggesting an average sales count for vehicles in this cluster.
Acceleration: The value is positive, implying higher than average acceleration for vehicles in this cluster.
TopSpeed: The value is negative, indicating a lower than average top speed for vehicles in this cluster.
















